
---
title: "WK7 Quiz"
course: Statistics & Data Analysis
output: html_document
Name:Pearl
---

Submission Format: Knit completed quiz to an HTML & upload via Canvas

## Scenario: CivicData - City Insights Lab

CivicData is a fictional analytics team working with several cities to monitor resident services, infrastructure efficiency, and response times. You have joined as a junior data consultant and have been asked to support ongoing work in three areas: housing inspections, bus wait times, and emergency service response.

Throughout the quiz, fill in the missing R code to complete simulations, summarize data, or visualize distributions. All code blocks require completion and interpretation. Run each of the code blocks upon completion.

---

### Question 1: Discrete vs. Continuous Data Types

The team collected the following information in the city of Kentmere:

- `inspections_completed`: number of housing inspections conducted each day (e.g., 12, 15, 17)
- `avg_wait_time`: average bus wait time at city stops in minutes (e.g., 4.2, 5.6, 6.1)

Identify which variable is discrete and which is continuous. Explain why classification matters when selecting probability distributions. Answer below:

inspections_completed is discrete. avg_wait_time is continuous. Classification matters when selecting probability distributions because affects how to graph and analyze the variables.  

Complete the code block to define sample vectors:

```{r}
inspections_completed <- c(12, 14, 13, 15, 17, 16)
avg_wait_time <- c(5.2, 4.8, 5.9, 6.3, 7.4, 7)
```



### Question 2: Simulating Discrete Events (Poisson)

The city of Auburn records the number of emergency calls received per hour. Historically, the average is 8 calls/hour. You are tasked with simulating this over 100 hours.

Complete the code to generate the distribution and summarize it:

```{r}
set.seed(101)
emergency_calls <- rpois(n = 100, lambda = 8)
table(emergency_calls)
```

What is the most common call count observed? How does this relate to the shape of a Poisson distribution? Answer below:

7 has the most common call count. The shape of a Poisson distribution is centered around its mean. The table shows how the calls have lower frequencies and the highest frequencies. 



### Question 3: Modeling Continuous Wait Times

You are asked to model bus stop wait times in the city of Larkpoint. Prior studies show that wait times follow a normal distribution with a mean of 6 minutes and a standard deviation of 1.2 minutes.

Complete the code to simulate and analyze 500 random wait times:

```{r}
set.seed(202)
wait_times <- rnorm(n = 500, mean = 6, sd = 1.2)
summary(wait_times)
mean(wait_times > 8)
```

  
Estimate the probability a rider waits more than 8 minutes. (Hint: Use `mean(wait_times > 8)`.) Answer below:

The probability a rider waits more than 8 minutes is 0.048.

### Question 4: Sampling Method Justification

The city of Norford is evaluating resident satisfaction by neighborhood income tier: low, middle, and high. Each tier has different population sizes.

Which sampling method would ensure all three income tiers are proportionally represented? Explain why your choice is appropriate. Answer below:

The sampling method I would use is stratified sampling. Stratified sampling is appropriate because we are studying smaller categories in a certain population. 

Complete the code structure for defining strata and generating a proportional sample:

```{r}
residents <- data.frame(
  income_tier = rep(c("Low", "Middle", "High"), times = c(300, 500, 200)),
  rating = sample(1:5, size = 1000, replace = TRUE)
)

library(dplyr)
stratified_sample <- residents %>%
  group_by(income_tier) %>%
  sample_frac(0.5)
```



### Question 5: Visualizing Distributions

The city of Bellacre wants to understand how water usage varies across homes. You collected water usage data (in gallons/day) for 1,000 homes.

Simulate this data using a normal distribution with mean = 180 and sd = 30, then plot a histogram and overlay a density line.

Complete the code below:

```{r}
set.seed(303)
water_use <- rnorm(n = 1000, mean = 180, sd = 30)

library(ggplot2)
ggplot(data.frame(water_use), aes(x = water_use, y = ..density..)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
  geom_density(color = "red", size = 1) +
  theme_minimal()
```

 
What does the shape of the distribution suggest? Is it symmetric, skewed, or irregular? Answer below:

Symmetric

---
