---
Title: Assignment 2
Name: Pearl Vang
---
Hypothesis: Increase in Unemployment rates correlates with lower House Price Index Scores
---
Load in the required libraries and dataset

```{r}
library(readxl)
library(Hmisc)
library(ggpubr)


data_set <- read_excel("C:\\Users\\Pearl\\OneDrive - Wayne State University\\Documents\\Coding\\Annual_Macroeconomic_Factors.xlsx")
```

Create a summary of the dataset for House Price Index and Unemployment Rates

```{r}
describe(data_set$House_Price_Index)
describe(data_set$Unemployment_Rate)

```

After research, t-test is not appropriate for analyzing the correlation between these two variables. I will use cor.test to test the relationship between the two variables and visualize it as a scatterplot.

```{r}
cor.test(data_set$House_Price_Index, data_set$Unemployment_Rate, method=c("pearson"))
ggscatter(data_set, x = "Unemployment_Rate", y = "House_Price_Index",
          add = "reg.line", conf.int = TRUE,
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Unemployment Rates", ylab = "House Price Index")

```